### 2023-12-07 analysis/new_analysis.qmd
new analysis

---
title: "Horse Racing Analysis"
author: "MC"
format:
  html:
    theme: flatly
    date: last-modified
    smooth-scroll: true
    cap-location: top
    code-fold: true
execute:
  warning: false
---

This is a quick attempt to (re-)analyse the horse-racing data from Badaya, Hartsuiker, Corley.  Prompted by the fact that Espe's original analysis (see [github.com/erbadaya/horse-race-task](https://github.com/erbadaya/horse-race-task)) produced a graph that looked like an interaction, but the analysis suggested that there were only main (or simple?) effects.

## Setup, Data Wrangling

All parts of this analysis are loosely based on Espe's code.  Here we load some relevant libraries, read in the experimental data, and get it broadly into shape.

```{r}
#| label: setup
library(tidyverse)
library(lme4)
library(here)

rawData <- here("data/horse_race_exp1_rawdata.csv")
df <- read_csv(rawData) |> select(-`...1`)

# function to ensure we have up-to-date bets for all horses

df_cond <- df |>
  filter (overbet=="no") |>
  select(ppt,horse,delivery,speaker) |>
  mutate(horseN=rep(c("horse0","horse1","horse2","horse3"),max(ppt)))

df_money <- df |> select(ppt,starts_with("horse")) |>
  group_by(ppt) |>
  filter(row_number()==1) |> select(-horse) |>
  pivot_longer(cols=starts_with("horse"),names_to="horseN",values_to="BET")

df_ready <- left_join(df_cond,df_money)
```
OK, so (see code above) `df_ready` is now a simple dataframe with the betting data.

We have to parse the questionnaire data, at least to take out people with dodgy audio(?) and also those highly familiar with horse-racing.  At the moment I'm not going to attempt a full analysis of the questionnaire data.  So to match Espe's criteria, we need to identify the questionnaire responses with "naturalness" in them, and extract scores for 3 values I think:  `naturalness_native`, `naturalness_nonnative`, and `P1_Q5` which is apparently betting expertise.

```{r}

rPatterns <- c(
  '.*P1_Q5":',
  gambling='\\d+',
  '.*naturalness_native":',
  natnat='\\d+',
  '.*naturalness_nonnative":',
  natnon='\\d+',
  '.*'
)

df_q <- df |> filter(str_detect(response,"country_native")) |>
  select(ppt,response) |>
  separate_wider_regex(response,patterns=rPatterns)

df_q <- df_q |> mutate(
  outAudio=ifelse(natnat<4 | natnon <4,"OUT","in"),
  outBetting=ifelse(gambling > 3, "OUT", "in")) |>
  select(-c(gambling,natnat,natnon))

```

Now we need to take out the 9 participants identified by Espe from comments:

```{r}
df_q <- df_q |> mutate(
  outVarious=case_when(
    ppt == 12 ~ "OUT", 
    ppt == 23 ~ "OUT",
    ppt == 40 ~ "OUT", 
    ppt == 46 ~ "OUT", 
    ppt == 279 ~ "OUT", 
    ppt == 334 ~ "OUT", 
    ppt == 505 ~ "OUT", 
    ppt == 586 ~ "OUT", 
    ppt == 617 ~ "OUT",  
    TRUE  ~ "in"
  )
)
```

> `r sum(df_q$outAudio=='OUT')` participants were excluded from
> the analysis because
> they detected that the audio had been manipulated.
> A further `r sum(df_q$outBetting=='OUT' & df_q$outAudio!='OUT')` were
> removed due to expertise in horse racing.
> 
> Finally, `r sum(df_q$outVarious=='OUT' & df_q$outBetting!='OUT' & df_q$outAudio!='OUT')` additional
> participants were removed because their feedback
> suggested that they had responded strategically during the
> study.

Last bit of data-munging: Remove participants from the "master" dataframe, and set contrasts for the experimental conditions.

```{r}
df_ready <- left_join(df_ready,df_q)
df_ready <- df_ready |> filter(outAudio=='in' &
                               outBetting=='in' &
                               outVarious=='in') |>
  select(-starts_with("out"))

df_ready <- df_ready |> mutate(delivery=factor(delivery,levels=c('fluent','disfluent')),
       speaker=factor(speaker)) # native comes before non
contrasts(df_ready$delivery) <- c(-.5,5)
contrasts(df_ready$speaker) <- c(-.5,5)
```

Finally, make one more column which is a "normalised bet" column (we act as if everyone had bet out of 100).

**In doing this, I noticed that (at least) one participant had bet zero on all horses, so let's remove them.**

```{r}
df_ready <- df_ready |> group_by(ppt) |>
  mutate(sum=sum(BET)) |> ungroup()

p_gone=sum(df_ready$sum==0)/4

df_ready <- df_ready |> filter(sum>0) |>
  select(-sum) |>
  group_by(ppt) |> mutate(BET100=BET*100/sum(BET))
```

> `r p_gone` remaining participants bet no money on any horse, and were removed
> from the analysis.

## Results!

Now draw a graph.  NB., we have two options here:  Either we scale up all of the bets so that they're "out of 100" or we leave as-is.  Starting with as-is:

```{r}
df_ready |> ggplot(aes(x=delivery,y=BET100,fill=speaker)) +
  geom_bar(position="dodge",stat="summary",fun.y="mean")
```