### 2024-09-09 manuscript/manuscript_eb.Rmd
write up & posters

---
title             : "Same as always: I suck at titles"
shorttitle        : "Horse race experiment"

author: 
  - name          : "Esperanza Badaya"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "esperanza.badaya@ugent.be"
  - name          : "Robert J. Hartsuiker"
    affiliation   : "1"
    - name          : "Martin Corley"
    affiliation   : "2"
affiliation:
  - id            : "1"
    institution   : "Department of Experimental Psychology, Ghent University, Ghent, Belgium."
  - id            : "2"
    institution   : "Department of Psychology, University of Edinburgh, Edinburgh, United Kingdom."

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Impressions of other people arise naturally and automatically [@ulemanetal2008]. Speakers can be evaluated not only by what they say but how they say it. For example, vocal features of speech can affect perceptions of confidence and persuasion [], status and solidarity [], and even attractiveness []. One repeately reported feature of speech said to bias listeners' evaluations is its fluency. Disfluencies, such as _uh_ or _um_ in English can lead to poorer evaluations of the speaker in terms of intelligence [], competence [], and certainty of their own knowledge []. Interestingly, speech produced with an accent different from one's own - particularly a _foreign_ accent - can also elicit negative perceptions of the speaker, including worse evaluations in terms of status, solidarity, and credibility (e.g., [], [], [], []). In this study, however, we explore a rather counterintuitive approach: That the co-presence of these two factors (i.e., (dis)fluency and foreign accents) can diminish the negative impact they have individually.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::